### 2024-04-17 paper/paper.qmd
added discussion points

---
title: "Precious Metals Or Crypto, Where Should I Invest?"
subtitle: "A Historical Analysis of the Two Major Asset Categories 2017-Present"
author: 
  - Yi Fei Pang
thanks: "Code and data supporting this analysis is available at: https://github.com/apang00/investment-comparison"
date: "`r Sys.time()`"
date-format: "D MMMM YYYY"
toc: true
format: 
  pdf:
    documentclass: article
    geometry: margin = 1in
    abstract: "This paper utilizes and compares datasets of the two most held precious metals and the two most held cryptocurrencies. A few simple price analyses were implemented to determine which assets were more profitable along with range formulas to determine volatility. Observations show that in the recent years, crypto has been a much more profitable investment. However, precious metals are much less volatile and consistent. The importance of this analysis can be very beneficial today due to the economic turmoil and investors asking: Do I want an asset that's profitable, consistent, or both?"
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(here)
library(knitr)
library(readr)
library(ggplot2)
```

{{< pagebreak >}}
# Introduction

Throughout history, human currency has always been evolving alongside our civilization. From the archaic days of the bartering systems where people traded raw materials, coins and precious metals emerged, then came paper money. As humans evolved even further, electronic banking with credit card emerged and now finally, crytpocurrencies. At every stage of society, humans have tied a value on an economic instrument as a way of wealth storage and wealth exchange.

Gold and silver has been an asset since the ancient civilizations of Egyptians and Incas. Both metals are malleable and therefore can easily be stamped into coins or molded into jewelry. Furthermore, pure gold and silver does not corrode or tarnish, thus making them timeless assets [@InvestopediaMetals]. Thousands of years later, in our modern era, gold and silver still remain as solid assets because of their durability, liquidity, and stability.Through economic turmoil, precious metals have proven to be an asset that not only holds it's value but has been significantly outperforming the stock markets since the turn of the millennium [@Forbes].In the recent decade, a new category of asset has emerged and performed tremendously well.That is the cryptocurrency. Originally developed as a method of making untraceable transactions and used almost exclusively by the Dark Web crowd, this technology has quickly gained popularity and has transformed into more of an investment rather than to obfuscate transactions. Cryptocurrencies such as Bitcoin and Ethereum has seen growths of ~20,000% and ~35,000% respectively at their peaks.

Since both precious metals and cryptocurrencies have been performing exceptionally well, this paper will analyse asset is a better hold depending on the goal of the investor. I will not only focus on the trends for the separate assets through price charts, but also evaluate correlations between the different category of assets through regression. Moreover, ranges will be calculated as a method to assess risk and volatility and compared between the two asset categories. Overall, it was evident precious metals tend to grow consistently over the year without being significantly affected by the economy or the introduction of other assets. In contrast, cryptocurrencies are much more volatile, seeing massive price ranges. However, cryptocurrencies tend to grow at much higher rates. Over the past decade, both precious metals and cryptocurrencies have been phenomenal assets and looks to continue to rise in value in the foreseeable future.

In the data section of the paper, there will be an introduction regarding where the data came from, and for what period the data was extracted. Furthermore, the data cleaning and visualization process will be detailed for creating the price, volatility, and regression charts. Next, the model section will explore the linear models used, why they were used, and how they represent the correlations. Following that, the results will be discussed includes elements such as the R-square values and p-values. To conclude, we will figure out potential weaknesses of the analysis along with the next steps and improvements.

{{< pagebreak >}}
# Data

## Data Description

This paper uses four separated datasets from two different sources. The gold  [@goldbusinessinsider] and silver [@silverbusinessinsider] raw datasets were sourced from a subsidiary of Business Insider, Market Insider. These two datasets include detailed financial information regarding gold and silver price and volume data starting from late 2017 until present. The other two cryptocurrency datasets for Ethereum [@ethyahoofinance] and Bitcoin [@btcyahoofinance] come from Yahoo Finance also ranging from late 2017 until present. All four datasets include details such as daily open and closing prices as well as high/lows and volume data. In addition, all prices in the data is in US dollars. The data is sourced from live charts for the daily open, close, and volume recorded on exchanges and recorded at the end of every day. The cleaning, modelling, and analysis of these four datasets will be carried out using the statistical programming language `R` [@citeR], using the packages `tidyverse`[@citetidy], `here`[@citeHere], `rstanarm`[@rstanarm], `dplyr` (Wickham et al. 2021), and `ggplot2`[@citeGG].

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(here)
library(knitr)

crypto <- read_csv(here::here("data/analysis_data/crypto prices.csv"))
metals <- read_csv(here::here("data/analysis_data/metals prices.csv"))
volatility <- read_csv(here::here("data/analysis_data/crypto volatility.csv"))
p_5 <- head(crypto, 5)
m_5 <- head(metals, 5)
v_5 <- head(volatility, 5)

```

## Price Data

The crypto prices and metals prices cleaned datasets feature the closing price points of the two cryptocurrencies eth_Close and btc_Close, along with the respective close prices for the metals gold_Close and silver_Close. In addition, there is a date column for the closing prices. These datasets range from November 12, 2017 until the date they were fetched, April 4th, 2024. Though there is data for every single day of the year within the aforementioned range, the cleaned dataset only takes into account the closing balance of the crytpocurrency/precious metal every 15 days, or the 1st and 15th of every month. The purpose is to reduce noise and filters out short term volatility and gives the data a more long term perspective. In addition to the date adjustment, there are adjustments made to the closing prices as well. For readability in graphs, all assets were readjusted to have a present value of in the vicinity of 1 Bitcoin. The adjustments made were 20x Ethereum, 30x Gold (oz.), and 2500x Silver (oz.).

```{r}
#| echo: false
#| warning: false
#| label: tbl-1
#| tbl-cap: "A summary table for crypto price data"

kable(p_5)
```

```{r}
#| echo: false
#| warning: false
#| label: tbl-2
#| tbl-cap: "A summary table for metal price data"

kable(m_5)
```

{{< pagebreak >}}
## Volatility Data

The volatility data was separated much like the prices datasets. One dataset for crypto volatility and one for metals volatility. In both datasets, volatility was calculated by using a normalized average true range.

First, the true range was calculated with:

TR = Max [(H - L), |H - Cp|, |L - Cp|]

Where:

H = Today's high

L = Today's low

Cp = Yesterday's closing price

Max = Highest value of the three terms

Source: [@InvestopediaFormula]

__________
Then the average true range was calculated with and represented by (asset)_ATR in @tbl-3:

(1/n) * Σ(TRi)

Where:
TRi = Particular true range for each period (e.g., first day's TR, second day's TR, etc.)

Σ = Summation symbol (summing up all the individual TRi values)

n = Number of periods (15 in the data to match the price data)

Source: [@InvestopediaFormula]

__________
In the final step, the normalized average true range was calculated by normalizing the ATR value with the closing price of the asset and represented by (asset)_ATR_norm in @tbl-3:

ATR/Cp

Where:

ATR = The ATR of the asset

Cp = Current day's closing price

{{< pagebreak >}}
## Data Visualization

```{r}
#| echo: false
#| warning: false
#| label: fig-1
#| fig-cap: "Price Comparison All Assets 2017 to Present"

compilation <- ggplot() +
  geom_line(data = crypto, aes(x = Date, y = eth_Close, color = "Ethereum"), size = 0.75) +
  geom_line(data = crypto, aes(x = Date, y = btc_Close, color = "Bitcoin"), size = 0.75) +
  geom_line(data = metals, aes(x = Date, y = gold_Close, color = "Gold"), size = 0.75) +
  geom_line(data = metals, aes(x = Date, y = silver_Close, color = "Silver"), size = 0.75) +
  labs(title = "Asset Prices",
       x = "Date", y = "Closing Amount ($USD)") +
  scale_color_manual(values = c("Ethereum" = "royalblue", "Bitcoin" = "sienna2", "Gold" = "#FFD700", "Silver" = "#C0C0C0"),
                     labels = c("Ethereum" = "Ethereum (20 Units)", "Bitcoin" = "Bitcoin", "Gold" = "Gold (30 oz.)", "Silver" = "Silver (2500 oz.)")) +
  guides(color = guide_legend(title = "Assets")) +  
  scale_x_date(date_breaks = "1 year", date_labels = "%Y")

compilation
```
The figure above shows the trends of all four separate assets. Over the past 7 years, it's evident that all the assets have increased in value significantly. Gold tends to have the least fluctuations and grew from close to 38,000 USD per 30 oz. to around 66,000 USD. Both precious metals are more stable than the cryptos. However, both ctyprocurrencies began at less than 10,000 (1 btc and 20 units of eth) and ended up well over 66,000 USD, resulting in a much more drastic return.

{{< pagebreak >}}
# Model
The goal of my model is to analyse the correlation between crytpo prices and precious metals prices. Through observing the previous price chart, there can be a perception of linear correlation between the two asset categories over 2017-2024 since all 4 assets show an increasing trend. Therefore, to represent this relationship, I will choose a linear regression model.

## Model set-up
The model will be linear where:

Define $y_b$ as the Bitcoin price and $y_e$ as the Ethereum price. Then $x_g$ and $x_s$ will be the price of gold and silver respectively.

$$
y_e + y_b =  b_0 + b_1(x_g + x_s) + c
$$
This can be simplified if the metals were aggregated into $y_i = y_e + y_b$ and $x_i = x_g + x_s$.

$$
y_i = b_0 + b_1x_i + c
$$
Where:

-   $y_i$ is the aggregated representation of the prices of the two major cryptocurrencies
-   $x_i$ is the aggregated representation of the prices of the two major precious metals
-   $b_0$ is the intercept where the price of the precious metals would be equal to zero
-   $b_1$ is the slope coefficient/estimated rate of change for cryptos for every one unit of change for metals
-   $c$ is the error variable that accounts for deviations from the predicted value

### Model Justification
I expect there to be a positive correlation between

# Results
```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <- readRDS(file = here::here("models/volatility chart.rds"))

```

## Price Results

## Volatility Results

Our results are summarized in...

```{r}
#| echo: false
#| warning: false
#| label: fig-3s
#| fig-cap: "Volatility Chart"

first_model
```
The Volatility chart shows that

{{< pagebreak >}}
# Discussion

## Precious Metals As An Investment
As the world enters a recession, we can confidently look at historical data for precious metals near and far and identify that it would do well to hold its value and steadily rise through economic failures and inflation. As shown in @fig-1, precious metals, especially gold is almost unaffected by the recession caused by Covid-19. In addition, @fig-2 demonstrates that the normalized average true range values for precious metals tend to be significantly lower than that of cryptocurrencies. That is suggestive that precious metals are stable investments and can be treated as very stable investments. As my previous models predict, precious metals will continue to increase in value in the future. As an investor, if you are looking to invest in precious metals the two most common options will be to either invest in derivative contracts or through gold mutual funds [@InvestopediaInv]. The second option would be to buy the physical metals in the form of coins or bars, however, this method is much more difficult to keep track of and lack mobility. Overall, precious metals is a great investment to dedicate to for those who wants to hedge against inflation, diversify their portfolio, while maintaining stability.

## Cryptocurrencies As An Investment
Cryptocurrencies is the new wave. Aside from Bitcoin and Ethereum, we are have seen the rise of many other successful cryptocurrencies such as Monero, Litecoin, and Stellar Lumens just to name a few. @fig-1 can attest to the substantial growth of cryptocurrencies over the last decade. However, @fig-2 shows that even though Bitcoin and Ethereum has been extremely profitable, they are both extremely volatile assets with normalized average true range values over 5 times that of precious metals. As an investor, cryptocurrencies can be much more appealing, especially towards the newer crowd. Investing in crypto is much more straightforward than precious metals. With large crypto exchanges such as Coinbase, Binance, and Kraken gaining popularity buying crypto has never been easier. Moreover, crypto wallets can come in the form of a USB Ledger or even an app on the appstore. This makes the asset extremely lightweight and mobile.

## Different Stroke For Different Folks
Both cryptocurrencies and precious metals are amazing investments.Throughout my analysis, I have concluded that there is a correlation between the two category of assets, both assets have seen massive growth, and both still have a lot of room to develop in the future. For the readers still asking: Which asset should I invest in? I believe the answer is this:

For those who are looking for something long term to hold onto. For those who want to sit on an asset for medium to long term, precious metals is the better investment. Whether you would want to buy futures, or invest in mutual funds, go for precious metals. This process will involve more time in effort whether it's purchasing or liquidating. Opening a mutual fund or dealing with derivative contracts is not simple and will take days to attain. Similarly, purchasing, carrying, and selling physical metals is also a tedious task. The result however, will be a long term investment that is stable and profitable.

For those who are looking for something long or short term, and that are more risk-seeking, cryptocurrencies will be your cup of tea. Though seemingly much more profitable in the recent years, these assets are much more volatile and is very dependent on economic climate. Moreover, compared to precious metals, buying, storing, and selling is a breeze. In cryptocurrencies is an asset with very high potential ceiling and as blockchain technology continues to progress, cryptocurrencies have a very bright future.

## Biases and Weaknesses
One of the greatest weaknesses stems from the limited historical data of cryptocurrencies. The value of precious metals can be trace back centuries and has solidified itself as a solid holding asset while cryptocurrencies such as Ethereum was only introduced in the late 2016s. Therefore, the only results and conclusion that can be made...

## Next Steps
Some next steps I can take for future developments can be to continue to monitor the price of crypto over the next few decades and see how it reacts to different economic conditions and how that compares to precious metals.

## Closing Remarks
Before investing in anything, nothing beats your own research. An investment should be considered only after sufficient self evaluation. Ask yourself:

-   How much money am I looking to invest?

-   What is my tolerance for risk?

-   What are my investment goals and objectives?

I have given but a brief overview an analyses of two popular modern day assets. The rest is up for you to discover.

\newpage

# References